Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-08-15 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in any localised nucleation events even while in the proximity of 
the droplet edge. During the experiments with the galvano-mirror, it was found 
that when no particle was present in the optical trap nucleation events would 
occur while the beam was close to the edge of the droplet, even though the solution
was undersaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but was more interesting is how the beam's motion influenced the growth of the nucleus.

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no real influence over the crystal shape, instead 
it grows outward from the trap focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the trap.}
	\label{fig:stationary_beam}
\end{figure}

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the necessary shear rate 
for crystal nucleation we wanted to see if a trapped silica bead could be trapped
and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was added to $10\ \mu L$ of a dilute water-silica mixture making the solution undersaturated ($S\approx0.7$). However, due to the beam's motion we instead 
encountered several unexpected results.

Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
moving in a small elliptical pattern; interestingly. After a few minutes we 
noted the appearance of several small droplets, these appear too small to be 
silica beads which have a uniform radius of $1.57\ \mu m$. We surmise these 
could be small clusters of Glycine that had previously been shown to form 
when the aqueous solutions where irradiated with a focused laser \cite{Tsuboi2009,
Gowayed2021}. While no droplets are seen directly entering the focus a nucleus 
forms close to the droplet edge, unlike in Fig.~\ref{fig:stationary_beam} the 
crystal does not grow out from the focal point evenly. Due to the galvano mirror, 
the crystal is simultaneously being moved by and growing around the focal point 
of the trap. Because of this the crystal nucleus lacks a clear morphology at 
first. Until roughly $20\ s$ the crystal reaches a almost prismatic structure, 
with further irradiation increasing the size.

Interestingly the galvano-mirror allows the trap to impart a slight torque 
on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
where even though the crystal is not directly in the trap focus it rotates 
in the $x-y$ plane and gets trapped again at a corner. The rotation could 
not be due to fluid flow close to the surface of the crystal as trap should 
have no influence on fluid molecules smaller than a few hundred nano-meters. 
In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal growth becomes 
localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth of a nucleus using 
		a moving beam. Initially the crystal shape is amorphous (a) but eventually
		reaches a more regular shape (b). This crystal is still influenced by the
		optical trap as even when not directly irradiated by the laser the crystal
		rotates between (c) and (d). When the laser is focused on a corner the 
		crystal growth is localised to that region, resulting in an elongated 
		section forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}. Though notably in both 
cases the crystal morphology is not influenced by the location of the laser focus, 
nor do they report the appearance of glycine droplets. Reports of these droplets
has been found only in supersaturated conditions, often using a lower NA objective
to provide a wider trap focus \cite{Liao2022, Yuyama2010, Yuyama2012}.

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought to the laser 
focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
again the beam was focused close to the droplet edge, this time the galvano mirror
was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
After a few minutes of irradiation droplets were seen entering the camera frame, 
because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping of a glycine 
		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
		black) being brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth is rather slow 
		with the crystal having no clear defined morphology through (c) and (d). 
		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates they must have a higher 
refractive index than the surrounding solution. However, getting an exact 
measurement of the

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation and particle displacement
did not result in any localised nucleation events even while in the proximity of 
the droplet edge. During the experiments with the galvano-mirror, it was found 
that when no particle was present in the optical trap nucleation events would 
occur while the beam was close to the edge of the droplet, even though the solution
was unsaturated. This has been reported prior \cite{Rungsimanon2010, Liao2022}, 
but was more interesting is how the beam's motion influenced the growth of the 
nucleus.

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no influence over the crystal shape, instead it 
grows outward from the trap focus. Furthermore, due to the fact that the solution
is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
the crystal escapes as its size exceeds the trap focus. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
	the trap.}
	\label{fig:stationary_beam}
\end{figure}

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the necessary shear rate 
for crystal nucleation we wanted to see if a trapped silica bead could be trapped
and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
encountered several unexpected results.

Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
moving in a small elliptical pattern; interestingly. After a few minutes we 
noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
for a larger example of these droplets). These appear too have a wide distribution
of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
\mu m$. We surmise these could be small clusters of Glycine that had previously 
been shown to form when the aqueous solutions where irradiated with a 
focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
directly entering the focus a nucleus forms close to the droplet edge, 
unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
being moved by and growing around the focal point of the trap. Because of 
this the crystal nucleus lacks a clear morphology at first. Until roughly 
$20\ s$ the crystal reaches a almost prismatic structure, with further 
irradiation increasing the size.

Interestingly the galvano-mirror allows the trap to impart a slight torque 
on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
where even though the crystal is not directly in the trap focus it rotates 
in the $x-y$ plane and gets trapped again at a corner. The rotation could 
not be due to fluid flow close to the surface of the crystal as the dipole 
moment of individual water molecules is too small to be influenced by an 
optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal 
growth becomes localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth of a nucleus using 
		a moving beam. Initially the crystal shape is amorphous (a) but eventually
		reaches a more regular shape (b). This crystal is still influenced by the
		optical trap as even when not directly irradiated by the laser the crystal
		rotates between (c) and (d). When the laser is focused on a corner the 
		crystal growth is localised to that region, resulting in an elongated 
		section forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
moving beam. This modification allows for the crystal growth to be localised to
a specific region of the bulk crystal whereas with a stationary beam there is no
control over the crystal morphology. In fact this allows for a much finer control
over the exact shape of the crystal nucleus, in some cases allowing for growth out
of the viewing plane as shown in

\begin{figure}
	\centering
\end{figure}

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought to the laser 
focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
again the beam was focused close to the droplet edge, this time the galvano mirror
was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
After a few minutes of irradiation droplets were seen entering the camera frame, 
because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping of a glycine 
		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
		black) being brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth is rather slow 
		with the crystal having no clear defined morphology through (c) and (d). 
		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates they must have a higher 
refractive index than the surrounding solution. It has been shown that the 
concentration of glycine solution is correlated with the refractive indices 
of the liquid \cite{Gowayed2021, Orttung1963}. This suggests that these droplets 
serve to provide solute material to the bulk crystal.

### 2024-08-15 Latex_files/Thesis/main_win.tex
updates
updates to ch3

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}
\usepackage{graphicx}
\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
\usepackage{sidecap}
% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

@@

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

@@

\title{PhD Title \\ PhD Thesis}
\author{D. Maciver\\[-0.8ex]
\small Chemical and Process Engineering Department\\[-0.8ex]
\small University of Strathclyde, Glasgow\\}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

>>>>

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
%\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}
\usepackage{graphicx}
\graphicspath{{C:/Users/xbb16146//dmaciver97/Latex_files/Thesis/figs}}
\usepackage{sidecap}
% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage[
backend=biber,
bibstyle=numeric,
citestyle=numeric,
minbibnames=3, maxbibnames=3,
]{biblatex}
\addbibresource{thesis_bib.bib}

@@

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{2}

@@

\title{Application of Micro-Rotors to Induce Crystal Nucleation: 
	   Challenges, Takeaways, and Future Prospects\\ PhD Thesis}
\author{D. Maciver\\[-0.8ex]
\small Chemical and Process Engineering Department\\[-0.8ex]
\small University of Strathclyde, Glasgow\\}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography
